---
title: "GWRpol"
author: "Linda Ilic"
date: "6 4 2022"
output: html_document
---
```{r}
#load libraries

easypackages::packages ("sf", "sp", "spdep", "spatialreg", "spgwr", "tmap", "mapview", "car", "RColorBrewer", "tidyverse", 
                        "cowplot", "leafsync", "leaflet.extras2", "mapview")

```

```{r}
#Before we do any analysis for GWR we need to convert our importanted sf object into a sp spatial object. As the spgwr package usually works on sp objects
#converting the sf polygon into sp object
#data <- st_read("data/Utrecht_2/output.shp")
data_sp <- as_Spatial(data)
#eq_poll <-  pm25_mean  ~ P_NW_MIG_A + P_LINK_HH + UITKMINAOW + WOZWONING +  GEM_HH_GR + HH_EENOUD + OAD + INW_65PL +INW_014 + GEBOORTE + P_WE_MIG_A + TOTHH_MPZK + WON_NBEW + WONVOOR45+WON_HCORP +WON_NBEW + WON_MRGEZ + P_HINK_HH +AV1_CAFE

eq_poll<-eq_pm25


#find the optimum bandwidth distance for fixed kernel using the gwr.sel() function from spgwr package
fbw_poll <- gwr.sel(eq_poll, 
               data = data_sp,
               longlat = TRUE,
               adapt=FALSE, 
               gweight = gwr.Gauss, 
               verbose = T)

fbw_poll
## [1] 5047.185
```

In our case, 5047.185 m is the optimum distance for the bandwidth. So we can use this to use the Fixed kernel for all 1673 Neighborhoods we have in our study area. So let us fit a fixed kernel GWR using the following code.


```{r}
fb_gwr_poll <- gwr(eq_poll, 
              data = data_sp,
              longlat = TRUE,
              bandwidth = fbw_poll, 
              gweight = gwr.Gauss,
              hatmatrix=TRUE, 
              se.fit=TRUE)

#summary of the model
fb_gwr_poll$results$rss


```
Let us skip the interpretation of this results immediately as these are not very helpful to get in idea about local variability of the results. What we are mostly interested in is about the local variability of the relationships and R2. To explore that we have to extract all the local models into a data frame and map the R2 and others. The next section of code do this works.

```{r}

#Extract the modeled relations for gwr object
fb_gwr_pout <- as.data.frame(fb_gwr_poll$SDF)

fb_gwr_pout

#see the data frame, each row is a regression result
view(fb_gwr_pout)


#join that with our main polygon data frame for the R2
data$fmb_localR2 <- fb_gwr_pout$localR2

mapview::mapview(data, zcol = "fmb_localR2", col.regions=brewer.pal(11, "RdYlGn"))


```

As you can see from the local R2, there is slight variations in R2 for each neighborhood. Same is true for coefficients. As we can also map each coefficients using the following codes.

```{r}
#join that with our main polygon data frame 
data$fcoef_GEM_HH_GR<-fb_gwr_pout$G_GAS_WON

coef_fb_gav <- mapview::mapview(data, zcol = "fcoef_GEM_HH_GR")

coef_fb_gav

```
Here, the effect of green space on premature mortality varies for different neighborhood. This local variations can tell, where the effects is more strong within the study regions. If you explore the maps, the dark purple shaded areas are the neighborhoods were higher green space has stronger negative correlations with lower premature mortality, compared to the yellow shaded neighborhoods. Such non-stationary variability in coefficients are really important to understand the local contexts and situation that maybe a function of the location itself! However, the fixed bandwidth approach may have limitations regarding the fixed distance applied for all the neighborhoods as data points search. This might be masking some key variability depending on the number of data points selected within fixed distance. To overcome this, we can apply more nuanced adaptive spatial kernl which is adaptive to the local settings.



5.2 Adaptive Kernel GWR: Selecting bandwidth, fitting GWR

Adaptive kernel do not depend on fixed distance as bandwidth. In this approch the kernel adapt in size to variations in the density of the spatial data so that the kernels have larger bandwidths where the data are sparse and have smaller bandwidths where the data are plentiful. Here, the bandwidth is not distance but numbers of nearest neighbors best fitted based on the data density. Here the k-nearest neighbors expressed as the proportion of observations (in this case the proportion of neighborhoods within at the neighborhoods in the data set). We can use the following code to find the optimal proportion as bandwidth.


```{r}
#adaptive kernel
abw_poll <- gwr.sel (eq_poll, 
              data = data_sp,
              adapt = TRUE, 
              gweight = gwr.Gauss)

#see the adaptive kernel 
abw_poll


```
Here the optimal bandwidth is 0.0083 indicating the proportion of observations (or number of neighbors) to be included in the weighting. In our case, we have 1673 neighborhoods in our data set, here 0.0083% of this total neighborhoods is equal to (1673 * 0.0083) ~ 13.88 or 14 nearest neighbors. For each neighborhood this approximate nearest neighborhoods should be used to calibrate the relevant local regression. Compared to the fixed kernel approch this is more adaptive to reflect the local surrounding conditions. Now we can use the following code to fit the adaptive kernel GWR.

```{r}

#Fitting the adaptive Kernel GWR
ab_gwr_poll <- gwr(eq_poll, 
              data = data_sp,
              longlat = TRUE,
              adapt = abw_poll, 
              gweight = gwr.Gauss,
              hatmatrix=TRUE, 
              se.fit=TRUE)

#summary of the model
ab_gwr_poll


```
Again the interpenetration of the printed result is not that important as they are not telling the local variability explicitly> Just have look at the Quasi-global R2, it is highest among all the previous models! We will now extract the local models into a data frame and map the R2 and others. The next section of code do this works.


```{r}
#adaptive GWR results in data frame
ab_gwr_pout <- as.data.frame(ab_gwr_poll$SDF)

#join the local R2 for each model to each neighborhood
data$amb_localR2 <- ab_gwr_pout$localR2

mapview::mapview(data, zcol = "amb_localR2", col.regions=brewer.pal(11, "RdYlGn")) 

```
Let us also check the variability of the coefficients for green space availability on premature mortality.

```{r}
#join that with our main polygon data frame 
data$acoef_G_GAS_WON = ab_gwr_pout$G_GAS_WON  

coef_ab_gav <- mapview::mapview(data, zcol = "acoef_G_GAS_WON")

coef_ab_gav

```
Here the size of the coefficient is larger then fixed kernel regression. Additionally we can see where there are strong negative relationship between green space exposure and premature mortality. Compared to global model with a single coefficient value here we can explore the local variability in more detail.

5.3 Compare the spatial variability of significance of relationships While we can explore the spatial variability of coefficients, but that does not tell us if the local relationships is statistically significant or not? Also what does it means in terms of interpretation? Some local relations maybe significant and some may be not, we can determine these by evaluating the t-statics for each indicator for each neighborhood. Here we can use the coefficient and standard error of the coefficient to estimate the absolute t value, and at a 0.05 significance lave t-values above 1.96 would indicate significant relationship between the dependent and independent variable. To note, there is another GWR package available that can also compute the p-value directly. The details of the package can be found here: Lu, B., Harris, P., Charlton, M. and Brunsdon, C., 2014. The GWmodel R package: further topics for exploring spatial heterogeneity using geographically weighted models. Geo-spatial Information Science, 17(2), pp.85-101. DOI: https://doi.org/10.1080/10095020.2014.917453

The following code can compute and map the statistical significance for both fixed and adaptive kernel models.


```{r}
#For fixed model
#estimate the t-value for ..... variable for fixed kernel model
data$ft_G_GAS_WON = fb_gwr_pout$G_GAS_WON / fb_gwr_pout$G_GAS_WON_se

#categorize the t-value to statistical significance
data$ft_G_GAS_WON_cat <- cut(data$ft_G_GAS_WON,
                             breaks=c(min(data$ft_G_GAS_WON), -1.96, 1.96, max(data$ft_G_GAS_WON)),
                             labels=c("sig","nonsig", "sig"))

#plot the significance for fixed kernel model
poll_sig_fb <- mapview::mapview (data, zcol = "ft_G_GAS_WON_cat")



#For Adaptive model
#estimate the t-value for variable for fixed kernel model
data$at_G_GAS_WON = ab_gwr_pout$G_GAS_WON / ab_gwr_pout$G_GAS_WON_se

#categorize the t-value to statistical significance
data$at_G_GAS_WON_cat <- cut(data$at_G_GAS_WON,
                             breaks=c(min(data$at_G_GAS_WON), -1.96, 1.96, max(data$at_G_GAS_WON)),
                             labels=c("sig","nonsig", "sig"))

#plot the significance for adaptive kernel model
poll_sig_ab <- mapview::mapview(data, zcol = "at_G_GAS_WON_cat")
poll_sig_ab
poll_sig_fb
#compare the maps
#poll_sig_ab | poll_sig_fb
```

